Thesis outline

### 2017-06-15 KellyDouglass/Thesis/Thesis_sections/GV.tex
Finished writing intro
Updates to Ch. 5-7

\chapter[Green valley galaxies]{Identifying green valley galaxies}

@@

\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    STATEMENT OF PROBLEM
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Chemical evolution and the green valley}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    BACKGROUND AND RELEVANCE TO PREVIOUS WORK
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{The color-magnitude diagram: Galaxies in the green valley}

@@

A galaxy's color is the ratio of the flux emitted by the galaxy in two different 
bands.  The stellar population of a galaxy will determine its color --- redder 
galaxies contain much older, cooler stars, while the spectrum of a bluer galaxy 
is produced by light from much younger, hotter stars.  A galaxy's color is 
independent of its redshift, so only photometry is needed to calculate the 
ratio.  While the KIAS-VAGC catalog (described below in Section \ref{sec:SDSS}) 
provides the $u-r$ color for all galaxies in SDSS, we make use of the Petrosian 
fluxes in the NSA (described in Section \ref{sec:GALEX}) to calculate the 
UV--optical color.  The NUV$-r$ color is calculated by
\begin{equation}
    \text{NUV}-r = -2.5\log \left( \frac{f_{NUV}}{f_r} \right)
\end{equation}
where $f_{NUV}$ is the Petrosian flux in the NUV-band of GALEX, and $f_r$ is the 
Petrosian flux in the $r$-band of SDSS.

@@

The SDSS Data Release 7 \citep[DR7]{Abazajian09} is a wide-field multi-band 
imaging and spectroscopic survey that uses drift scanning to map approximately 
one-quarter of the northern sky.  A dedicated 2.5-meter telescope at the Apache 
Point Observatory in New Mexico takes photometric data in the five-band SDSS 
system --- $u$, $g$, $r$, $i$, and $z$ \citep{Fukugita96,Gunn98}.  Galaxies with 
a Petrosian $r$-band apparent magnitude $m_r < 17.77$ are selected for 
spectroscopic analysis \citep{Lupton01,Strauss02}.  Gas-phase chemical 
abundances are calculated via the Direct $T_e$ method using emission-line fluxes 
measured by the MPA-JHU catalog\footnote{Available at 
\url{http://www.mpa-garching.mpg.de/SDSS/DR7}} and published in Douglass \& 
Vogeley (2017, in prep).  The large-scale environment of the galaxies is based 
on the void catalog compiled by \cite{Pan12}.  This void catalog is constructed 
from SDSS DR7 using the VoidFinder algorithm of \cite{Hoyle02}.  Galaxies which 
are located within the identified voids are labeled as void galaxies; those 
which do not reside in a void are designated as wall galaxies.  Because the 
minimum diameter of a void is defined as 10 \hMpc , the large-scale environment 
of galaxies within 5 \hMpc of the edge of the SDSS DR7 footprint cannot be 
described, so their environment is labeled as unknown.

% KIAS-VAGC (morphological classification)
The Korea Institute for Advanced Study Value-Added Galaxy Catalog 
\citep[KIAS-VAGC]{Choi10} contains galaxies from the SDSS DR7 main sky survey 
based on the New York University Value-Added Galaxy Catalog Data Release 7 
\citep[NYU-VAGC]{Blanton05}.  It provides a morphological class and type 
following the automated morphology classification scheme of \cite{Park05}, using 
the color $u-r$, color gradient $\Delta (g-r)$, and inverse concentration index.  
We use the absolute magnitudes as listed in the KIAS-VAGC.

\subsection{UV data --- GALEX}\label{sec:GALEX}
The Galaxy Evolution Explorer \citep[GALEX]{Martin05} is an orbiting space 
telescope conducting an extra-galactic ultraviolet all-sky survey.  The 
instrument allows imaging and spectroscopic observations in two ultraviolet 
bands --- Far UV (FUV, 1350--1780\AA) and Near UV (NUV, 1770--2730\AA) 
\citep{Morrissey07}.  The NUV-$r$ colors used in this study are calculated from 
the azimuthally-averaged SDSS-style Petrosian fluxes provided in the NASA-Sloan 
Atlas version 0.1.2 \citep[NSA]{Blanton11}.

@@

This analysis reveals a population of galaxies that lie in the Green Valley on 
the NUV$-r$ color-magnitude diagram, identified independent of their location on 
the diagram.  Rather than being classified based on arbitrary limits placed on 
the NUV$-r$ values \citep[as done by][]{Schawinski14,Salim14a}, the galaxies can 
be easily separated into three distinct populations based on the galaxies' 
morphological types as calculated in the KIAS-VAGC; this can be seen in Fig. 
\ref{fig:NUVr_CMD}.  Defined by the $u-r$ color and the color gradient 
$\Delta (g-i)$, it is novel that an estimate of the morphological type of the 
galaxies is an adequate classification for the three populations of the NUV$-r$ 
CMD.  In addition, Fig. \ref{fig:NUVr_hist} shows a histogram of the galaxies 
which overlap in both GALEX and SDSS, separated by their morphological 
classification from the KIAS-VAGC.  It shows that the three populations are well 
fit by three Gaussians --- the new population in the middle is the transient 
galaxy population originally ``missing'' from the optical CMD.

@@

If green valley galaxies are transitioning between the blue sequence and red 
cloud, then they should have intermediate (s)SFRs.  The distribution of SFR and 
sSFR for the galaxies in SDSS DR7 are shown in Fig. \ref{fig:SFR_hist}, with the 
galaxies separated by their morphological types.  We see that the distribution 
of SFRs for galaxies in the green valley peaks around 
$\log(\text{SFR}) \sim -0.5$, in between the ranges occupied by the red cloud 
and blue sequence galaxies.  A similar distribution is shown for the sSFRs, 
although the green valley galaxies extend well into the range of sSFRs occupied 
by the galaxies in the red cloud.  \cite{Salim14a} describes transitional 
galaxies (those in the green valley) to have lower sSFRs than galaxies in the 
blue sequence, specifically $-11.8 < \log(\text{sSFR}) < -10.8$.  While our 
results do indicate that green valley galaxies have lower sSFRs than those in 
the blue sequence, we find that their range of sSFR is almost double the width 
of that described in \cite{Salim14a}.  The green valley galaxies have a sSFR 
within the range $-12 \lesssim \log(\text{sSFR}) \lesssim -10$.

@@

Estimates of the gas-phase chemical abundances of galaxies probe their 
integrated star formation histories, which can provide insight into the 
enrichment of the ISM and CGM.  A few of the galaxies for which Douglass \& 
Vogeley (2017, in prep) are able to estimate metallicities reside in the green 
valley.  When compared with the metallicities of galaxies in the blue sequence 
and red cloud, Fig. \ref{fig:Z_hist} shows that the green valley galaxies have 
higher nitrogen abundances than star-forming galaxies in the blue sequence (and 
some of the red cloud).  With oxygen abundances that fall within the same range 
as the other two galaxy populations, this shift causes the N/O ratio to be much 
higher in green valley galaxies than in the blue sequence.

@@

\section[Discussion]{Understanding green valley galaxies}

@@

As explained in \cite{Douglass17b}, nitrogen is thought to be produced in both a 
primary and secondary stage, depending on the presence of other heavy elements.  
The fact that the galaxies in the green valley have higher nitrogen abundances 
compared to those in the blue sequence indicate that they are building up a 
surplus of nitrogen, most likely due to the presence of heavier elements at the 
later star formation episodes.  The surplus of heavier elements allows the CNO 
cycle to commence earlier in the stars' lifetimes, producing nitrogen at a 
higher rate than oxygen.  It is also possible that the higher nitrogen 
abundances are a sign that the galaxies' intermediate mass stars (those 
primarily responsible for nitrogen synthesis) have expired.  This surplus of 
nitrogen results in the high N/O ratios seen for the green valley galaxies in 
Fig. \ref{fig:Z_hist}.

Fig. \ref{fig:Mr_bin} shows a higher fraction of faint wall galaxies than faint 
void galaxies are in the green valley.  This would indicate that void galaxies 
are slightly behind wall galaxies in their evolution, matching predictions based 
on the $\Lambda$CDM cosmology.  These results also coincide with the 
observations of \cite{Douglass17b} and Douglass \& Vogeley (2017, in prep).  
They find that, for dwarf galaxies ($M_r > -17$), those which exist in voids may 
have commenced their star formation at a later time and have not experienced the 
same cosmic downsizing as dwarf galaxies residing in denser regions.  The higher 
fraction of wall dwarf galaxies in the green valley might also point to an 
environmental influence on the reason for transitioning through the green 
valley, such as ram pressure stripping, which would occur much less frequently 
in the void environment.  However, the existence of void galaxies in the green 
valley indicate that galaxy interactions cannot be the only star formation 
quenching mechanism.

@@

To understand the gas-phase chemical abundances of the green valley galaxies, we 
compare Fig. \ref{fig:Z_hist} with the distribution of the gas-phase chemical 
abundances of star-forming and composite galaxies and galaxies with an AGN shown 
in Fig. \ref{fig:BPT_abund}.  We see that the galaxies with an AGN have the most 
similar distributions in the three abundance ratios as the green valley 
galaxies: both groups have similar oxygen abundances (O/H) and higher nitrogen 
abundances (N/H) and N/O ratios than the star-forming galaxies.  This might 
indicate a link between green valley galaxies and a host AGN.  The presence of 
an AGN in green valley galaxies would support other studies of the transition of 
galaxies from the blue sequence to the red cloud, which attribute the quenching 
of star formation to accretion of a central black hole \citep[e.g.,][]{Croton06,
Stasinska08}.

@@

We can define galaxies which are transitioning through the green valley of the 
color-magnitude diagram by combining the galaxy's color, color gradient, and 
inverse concentration index.  Galaxies with a magnitude less than 25 for the 
morphological type as calculated in the KIAS-VAGC exist in the green valley 
portion of the UV-optical color magnitude diagram.  Galaxies defined as early 
types (morphological type values equal to 1 or 2) are in the red cloud, while 
those with morphological type values greater than 25 are in the blue sequence.  
With this quantitative definition for galaxies transitioning through the green 
valley, we can begin to understand the properties of green valley galaxies and 
the evolution of galaxies.

@@

There is a higher fraction of faint wall galaxies than faint void galaxies in 
the green valley.  In conjunction with the conclusions of \cite{Douglass17b} and 
Douglass \& Vogeley (2017, in prep), this indicates that void dwarf galaxies are 
less evolved than dwarf galaxies in denser environments.  It would be beneficial 
to investigate the small-scale ($\sim 1$ \hMpc) environment of galaxies in the 
green valley, to determine if any interactions are responsible for the 
transitional state.

>>>>

\chapter[Green valley galaxies]{Identifying green valley galaxies}\label{ch:GV}

@@

\section{Introduction}

@@

A galaxy's color is a logarithmic ratio of the ratio of the flux emitted by the 
galaxy in two different bands.  The stellar population of a galaxy will 
determine its color: redder galaxies contain much older, cooler stars, while the 
spectrum of a bluer galaxy is produced by light from much younger, hotter stars.  
While the KIAS-VAGC catalog (described below in Section \ref{sec:SDSS}) provides 
the $u-r$ color for all galaxies in SDSS, we make use of the Petrosian fluxes in 
the NSA (described in Section \ref{sec:GALEX}) to calculate the UV--optical 
color.  The NUV$-r$ color is calculated by
\begin{equation}
    \text{NUV}-r = -2.5\log \left( \frac{f_{NUV}}{f_r} \right)
\end{equation}
where $f_{NUV}$ is the Petrosian flux in the NUV-band of GALEX, and $f_r$ is the 
Petrosian flux in the $r$-band of SDSS.

@@

The SDSS Data Release 7 \citep[DR7;][]{Abazajian09} is a wide-field multi-band 
imaging and spectroscopic survey that uses drift scanning to map approximately 
one-quarter of the northern sky.  A dedicated 2.5-meter telescope at the Apache 
Point Observatory in New Mexico takes photometric data in the five-band SDSS 
system --- $u$, $g$, $r$, $i$, and $z$ \citep{Fukugita96,Gunn98}.  Galaxies with 
a Petrosian $r$-band apparent magnitude $m_r < 17.77$ are selected for 
spectroscopic analysis \citep{Lupton01,Strauss02}.  Gas-phase chemical 
abundances are calculated via the Direct $T_e$ method using emission-line fluxes 
measured by the MPA-JHU catalog\footnote{Available at 
\url{http://www.mpa-garching.mpg.de/SDSS/DR7}} and published in Douglass et al. 
(2017, in prep).  The large-scale environment of the galaxies is based on the 
void catalog compiled by \cite{Pan12}.  This void catalog is constructed from 
SDSS DR7 using the VoidFinder algorithm of \cite{Hoyle02}.  Galaxies which are 
located within the identified voids are labeled as void galaxies; those which do 
not reside in a void are designated as wall galaxies.  Because the minimum 
diameter of a void is defined as 10 \hMpc , the large-scale environment of 
galaxies within 5 \hMpc of the edge of the SDSS DR7 footprint cannot be 
described, so their environment is labeled as unknown.

% KIAS-VAGC (morphological classification)
The Korea Institute for Advanced Study Value-Added Galaxy Catalog 
\citep[KIAS-VAGC;][]{Choi10} contains galaxies from the SDSS DR7 main sky survey 
based on the New York University Value-Added Galaxy Catalog Data Release 7 
\citep[NYU-VAGC;][]{Blanton05}.  It provides a morphological class and type 
following the automated morphology classification scheme of \cite{Park05}, using 
the color $u-r$, color gradient $\Delta (g-r)$, and inverse concentration index.  
With these three quantities, galaxies are separated into one of three different 
morphological types: normal late type, normal early type, blue early type, and 
one of two different morphological classes: elliptical or lenticular and spiral 
or irregular.  For reference, typical star-forming spiral galaxies would be 
classified as a normal late type spiral or irregular galaxy, star-forming dwarf 
galaxies are usually normal late type spiral or irregular galaxies, and red 
elliptical galaxies are typically classified as normal early type elliptical or 
lenticular galaxies.  We use the absolute magnitudes as listed in the KIAS-VAGC.

\subsection{UV data --- GALEX}\label{sec:GALEX}
The Galaxy Evolution Explorer \citep[GALEX]{Martin05} is an orbiting space 
telescope conducting an extra-galactic ultraviolet all-sky survey.  The 
instrument allows imaging and spectroscopic observations in two ultraviolet 
bands: Far UV (FUV, 1350--1780\AA) and Near UV (NUV, 1770--2730\AA) 
\citep{Morrissey07}.  The NUV-$r$ colors used in this study are calculated from 
the azimuthally-averaged SDSS-style Petrosian fluxes provided in the NASA-Sloan 
Atlas version 0.1.2 \citep[NSA]{Blanton11}.

@@

This analysis reveals a population of galaxies that lie in the Green Valley on 
the NUV$-r$ color-magnitude diagram, identified independent of their location on 
the diagram.  Rather than being classified based on arbitrary limits placed on 
the NUV$-r$ values \citep[as done by][]{Schawinski14,Salim14a}, the galaxies can 
be easily separated into three distinct populations based on the galaxies' 
morphological types as calculated in the KIAS-VAGC; this can be seen in Fig. 
\ref{fig:NUVr_CMD}.  Here, we group together normal late type and blue early 
type galaxies, and we separate into two groups the normal late type galaxies 
based on their value of the morphological type quantity calculated in the 
KIAS-VAGC.  Defined by the $u-r$ color and the color gradient $\Delta (g-i)$, it 
is novel that an estimate of the morphological type of the galaxies is an 
adequate classification for the three populations of the NUV$-r$ CMD.  In 
addition, Fig. \ref{fig:NUVr_hist} shows a histogram of the galaxies which 
overlap in both GALEX and SDSS, separated by their morphological classification 
from the KIAS-VAGC.  It shows that the three populations are well fit by three 
Gaussians --- the new population in the middle is the transient galaxy 
population originally ``missing'' from the optical CMD.

@@

If green valley galaxies are transitioning between the blue sequence and red 
cloud, then they should have intermediate (s)SFRs.  The distribution of SFR and 
sSFR for the galaxies in SDSS DR7 are shown in Fig. \ref{fig:SFR_hist}, with the 
galaxies separated by their morphological types.  We see that the distribution 
of SFRs for galaxies in the green valley peaks around 
$\log(\text{SFR}) \sim -0.5$, in between the ranges occupied by the red cloud 
and blue sequence galaxies.  A similar distribution is shown for the sSFRs, 
although the green valley galaxies extend well into the range of sSFRs occupied 
by the galaxies in the red cloud.  \cite{Salim14a} finds transitional galaxies 
(those in the green valley) to have lower sSFRs than galaxies in the blue 
sequence, specifically $-11.8 < \log(\text{sSFR}) < -10.8$.  While our results 
do indicate that green valley galaxies have lower sSFRs than those in the blue 
sequence, we find that their range of sSFR is almost double the width of that 
described in \cite{Salim14a}.  The green valley galaxies have a sSFR within the 
range $-12 \lesssim \log(\text{sSFR}) \lesssim -10$.

@@

Estimates of the gas-phase chemical abundances of galaxies probe their 
integrated star formation histories, which can provide insight into the 
enrichment of the ISM and CGM.  A few of the galaxies for which Douglass et al. 
(2017, in prep) are able to estimate metallicities reside in the green valley.  
When compared with the metallicities of galaxies in the blue sequence and red 
cloud, Fig. \ref{fig:Z_hist} shows that the green valley galaxies have higher 
nitrogen abundances than star-forming galaxies in the blue sequence (and some of 
the red cloud).  With oxygen abundances that fall within the same range as the 
other two galaxy populations, this shift causes the N/O ratio to be much higher 
in green valley galaxies than in the blue sequence.  Our interpretation of this 
shift is described in Sec. \ref{sec:discussion_GV}.

@@

\section[Discussion]{Understanding green valley galaxies}\label{sec:discussion_GV}

@@

As explained in \cite{Douglass17b}, nitrogen is thought to be produced in both a 
primary and secondary stage, depending on the presence of other heavy elements.  
The fact that the galaxies in the green valley have higher nitrogen abundances 
compared to those in the blue sequence indicate that they are building up a 
surplus of nitrogen, most likely due to the presence of heavier elements in 
later episodes of star formation.  The surplus of heavier elements allows the 
CNO cycle to commence earlier in the stars' lifetimes, producing nitrogen at a 
higher rate than oxygen.  It is also possible that the higher nitrogen 
abundances are a sign that the galaxies' intermediate mass stars (those 
primarily responsible for nitrogen synthesis) have expired.  This surplus of 
nitrogen results in the high N/O ratios seen for the green valley galaxies in 
Fig. \ref{fig:Z_hist}.

Fig. \ref{fig:Mr_bin} shows a higher fraction of faint wall galaxies than faint 
void galaxies are in the green valley.  This would indicate that void galaxies 
are slightly behind wall galaxies in their evolution, matching predictions based 
on the $\Lambda$CDM cosmology.  These results also coincide with the 
observations of \cite{Douglass17b} and Douglass \& Vogeley (2017, in prep).  
They find that, for dwarf galaxies ($M_r > -17$), those which exist in voids may 
have commenced their star formation at a later time.  The higher fraction of 
wall dwarf galaxies in the green valley might also point to an environmental 
influence on the reason for transitioning through the green valley, such as ram 
pressure stripping, which would occur much less frequently in the void 
environment.  However, the existence of void galaxies in the green valley 
indicate that galaxy interactions cannot be the only star formation quenching 
mechanism.

@@

Galaxies can be classified into a number of categories with the use of a BPT 
diagram \citep{Baldwin81}.  By comparing different emission line ratios to get 
an idea of the hardness of the spectrum, we can identify if most of the energy 
in the galaxy's spectrum is coming from a star-forming region, an AGN, or a 
combination of both (``composite'').  To understand the gas-phase chemical 
abundances of the green valley galaxies, we compare Fig. \ref{fig:Z_hist} with 
the distribution of the gas-phase chemical abundances of star-forming and 
composite galaxies and galaxies with an AGN shown in Fig. \ref{fig:BPT_abund}; 
the BPT diagram classifications are from \cite{Brinchmann04}.  We see that the 
galaxies with an AGN have the most similar distributions in the three abundance 
ratios as the green valley galaxies: both groups have similar oxygen abundances 
(O/H) and higher nitrogen abundances (N/H) and N/O ratios than the star-forming 
galaxies.  This might indicate a link between green valley galaxies and a host 
AGN.  The presence of an AGN in green valley galaxies would support other 
studies of the transition of galaxies from the blue sequence to the red cloud, 
which attribute the quenching of star formation to accretion of a central black 
hole \citep[e.g.,][]{Croton06,Stasinska08}.

@@

We find a remarkable conformity between galaxies identified by morphological 
classification and their location in the NUV$-r$ CMD, which clearly identifies 
the green valley.  We can define galaxies which are transitioning through the 
green valley of the color-magnitude diagram by combining the galaxy's color, 
color gradient, and inverse concentration index.  Using these properties for the 
morphological classification, galaxies with a value less than 25 for the 
morphological type as calculated in the KIAS-VAGC exist in the green valley 
portion of the UV-optical color magnitude diagram.  Galaxies defined as early 
types (morphological type values equal to 1 or 2) are in the red cloud, while 
those with morphological type values greater than 25 are in the blue sequence.  
With this quantitative definition for galaxies transitioning through the green 
valley, we can begin to understand the properties of green valley galaxies and 
the evolution of galaxies.

@@

There is a higher fraction of faint wall galaxies than faint void galaxies in 
the green valley.  In conjunction with the conclusions of \cite{Douglass17b} and 
Douglass et al. (2017, in prep), this indicates that void dwarf galaxies are 
less evolved than dwarf galaxies in denser environments.  It would be beneficial 
to investigate the small-scale ($\sim$1 \hMpc) environment of galaxies in the 
green valley, to determine if any interactions are responsible for the 
transitional state.
